---
title: "Week 3 Assignment: Create New Choose Function"
author: "Melissa Childs"
date: "July 30, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For this week's assignment, we were tasked with creating our own version of the R choose() function to solve for "5 choose 3". The constraints are that we cannot use the built in factorial() and choose() functions.

In evaluating the equation for n choose r, which is "n!/((n-r)!*r!)", I first needed to create three different factoral functions, outlined below:

```{r}
#factorialn <- calculates n!
#factorialr <- calculates r!
#factorialnr <- calculates (n-r)!
```

Then, I needed to create a fourth equation, which would incoporate the three factorial functions into the choose equation:
```{r}
# mychoosee <- factorialn / factorialnr * factorialr
```
Then, I needed to incorporate all of this into a function, which would allow any value for n and r to be input into the mychoose and factorial equations and the correct solution for "n choose r" to be output.

So first, I created the factorial functions. I tried incorporating the factorial for-loop that I had developed in week 1 by assigning this loop to a function, and moving the "print" statement outside of the loop curly brackets so as to only print the final number of the factorial instead of the entire string.

```{r}
example <- function(x){
  myfact <- 1
for(i in 1:x)
{
 myfact <- myfact*i
 
}
 print(myfact)
}
example(12)
```

This was a sufficient solution for creating three unique factorial functions. However, I found that I was having a difficult time wrapping all the code for the loops into a fourth function for the choose equation.

This is when I realized that I needed a more simple function for factorials. I searched online, and found inventive ways to code for factorials on Matlab (cite: http://www.matrixlab-examples.com/factorials.html). This gave me the idea to use the built in "prod" function to calculate the factorials. 

The prod function works by multiplying together all the values included in its arguments. For example, using the prod function on a vector of numbers 1-5 will multiple each number in the sequence together:
```{r}
prod(1:5)
#we can check this value against the factorial function and see that it's correct, which it is
factorial(5)
```
So by replacing the value "5" in the prod argument with the value of n or r, and then assigning the built-in prod function to a variable, I can create a custom factorial equation that is equal to the prod function of 1 through n, r, or n-r.
```{r}
#factorialn <- prod(1:n)
#factorialr <- prod(1:r)
#factorialnr <- prod(1:(n - r))
```
Now I can incorporate these factorial functions into the choose equation:
```{r}
#mychoose2 <- factorialn / (factorialnr*factorialr)
# 
```
The final thing I needed to do was wrap all these equations into a function so that n and r are defined as arguments in the function before being input through the equations. My final code is below:
```{r}
mychoose <- function(n, r){
  factorialn <- prod(1:n)
  factorialr <- prod(1:r)
  factorialnr <- prod(1:(n - r))
  mychoose2 <- factorialn / (factorialnr*factorialr)
  return(mychoose2)
}
#Now all I have to do is input the values for n and r into the mychoose function 
mychoose(5, 3)
#we can compare this with the real choose function to confirm the answer
choose(5, 3)
```
